---
title: "Explaining iscambinompower"
author: "Shannon Pileggi"
date: "April 3, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r}
# Source function
source('~/Senior Project/SP--Pablo--RProgramming/ISCAM2/R/iscambinompower.R', echo=TRUE)
```

# Execute function

```{r, fig.height = 8}
iscambinompower(LOS = 0.05, 
                n = 20,
                prob1 = 0.25,
                alternative = "greater",
                prob2 = 0.333)
```

# Explain results

This execution has

+ $\alpha = 0.05$

+ $n = 20$

+ $H_0$: $p = 0.25$ vs $H_a$: $p > 0.25$, where $p = 0.333$

The top graph shows the *rejection region*:  that we would need to observe at least 9 successes in 20 trials in order to provide convincing evidence in favor of $H_a$, that $p$ exceeds 0.25.  If we observe 9 or more successes in 20 trials, our $p$-value for this test is 0.04093.  (This computation **does not** depend on the value of $p = 0.333$.)

The bottom graph shows the *power* of the test.  From the top figure, we already know that we would reject the null hypothesis if we observe at least 9 successes in 20 trials.  Given that that $p = 0.333$, we now know that the probability that we would observe at least 9 successes in 20 trials (i.e., that we would correctly reject the null hypothesis) is 0.1897.  (This computation **does** depend on the value of $p = 0.333$.)

# Demo rejection region calculation

```{r}
options(scipen=999) # this option prevents scientific notation
check <- round(data.frame(x = 0:20, 
                    prob1 = dbinom(0:20, 20, 0.25), 
                    cum_prob1 = pbinom(0:20, 20, 0.25, lower.tail = F),
                    prob2 = dbinom(0:20, 20, 0.33), 
                    cum_prob2 = pbinom(0:20, 20, 0.33, lower.tail = F)), 5)
check
```

