---
title: "ISCAM Package"
author: "Noelle Pablo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ISCAM Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package contains a variety of R functions to accompany the introductory statistics textbook, _Investigating Statistiscal Concepts, Applications, and Methods_ (ISCAM), written by Dr. Allan Rossman and Dr. Beth Chance. These functions are to aid students in their investigations, assignments, and projects. The functions are categorized into 5 groups: Data Visualization, Summary Statistics, Hypothesis Testing, Quantiles, Rejection Regions & Power, Probability Distributions, Normal Approximations, and Overlaying Probability Distributions.

To install the ISCAM package, run the following code:
```{r, message = FALSE}
install.packages("devtools") # If you already have the devtools package installed, you may skip this line
library(devtools)
devtools::install_github("shannonpileggi/SP--Pablo--RProgramming/ISCAM", build_vignettes = TRUE)
library(ISCAM)
```
## Data Visualization
- **iscam_boxplot**

This function displays horizontal boxplot(s). 
```{r, fig.width=6, fig.height=6}
data("chickwts")
iscam_boxplot(x = chickwts$weight, explanatory = chickwts$feed, names = c("Weight", "Feed"))
```

- **iscam_dotplot**

This function displays horizontal dotplot(s). 
Optional: A second, categorical variable can also be specified and values will be calculated separately for each group. 
 
```{r}
iscam_dotplot(response = chickwts$weight, explanatory = chickwts$feed, names = c("Weight", "Feed"))
```
 
## Summary Statistics
- **iscam_summary**

This function calculates the five number summary, mean, and standard deviation of the quantitative variable x.

```{r}
iscam_summary(x = chickwts$weight)
```
## Hypothesis Testing
- **iscam_onepropztest**

This function calculates a one proportion z-test and/or interval.
```{r, fig.width=6, fig.height=4}
iscam_onepropztest(observed = 12, n = 15, hypothesized = .5, alternative = "two.sided", conf.level = 0.95)
```

- **iscam_twopropztest**

This function calculates a two proportion z-test and/or interval. 

```{r, fig.width=6, fig.height=4}
iscam_twopropztest(observed1 = 6, n1 = 17, observed2 = 12, n2 = 20, hypothesized = 0, alternative = "less")
```

- **iscam_onesamplet**

This function calculates a one sample t-test and/or interval from summary statistics.

```{r, fig.width=6, fig.height=4}
iscam_onesamplet(xbar = 70, sd = 4, n = 25, hypothesized = 68, alternative = "greater")
```

- **iscam_twosamplet**

This function calculates a two sample t-test and/or interval from summary data. 

```{r, fig.width=6, fig.height=4}
iscam_twosamplet(x1 = 97.25, sd1 = 3.65, n1 = 8, x2 = 87.25, sd2 = 9.60, n2 = 12, hypothesized = 0, alternative = "less", conf.level = 0.95)
```

- **iscam_binomtest**

This function performs an exact binomial test and graphs the binomial distribution and/or binomial confidence interval.

```{r, fig.width=6, fig.height=4}
iscam_binomtest(10, 40, hypothesized = .50, alternative = "two.sided", conf.level = 0.90)
```

## Quantiles
- **iscam_invbinom**

This function calculates the binomial quantile of a specified probability. The integer that achieves at most the stated probability will be returned.

```{r, fig.width=6, fig.height=4}
iscam_invbinom(alpha = .05, n = 15, prob = .20, lower.tail = FALSE)
```

- **iscam_invnorm**

This function calculates the normal quantile of a specified probability. 

```{r, fig.width=6, fig.height=4}
iscam_invnorm(prob1 = .55, mean = 15, sd = 2, direction = "below")
```

- **iscam_invt**

This function calculates the t quantile of a specified probability.

```{r, fig.width=6, fig.height=4}
iscam_invt(prob1 = .35, df = 2, direction = "above")
```

## Rejection Regions & Power
- **iscam_binompower**

This function determines the rejection region corresponding to the level of significance and the first probability. A second, optional probability can be specified to determine the power of the test. 

```{r, fig.width=6, fig.height=6}
iscam_binompower(LOS = .05, n = 20, prob1 = 0.25, alternative = "greater", prob2 = 0.333, explain = T)
```

- **iscam_normpower**

This function determines the rejection region corresponding to the level of significance and the first probability and shows the second distribution shading its corresponding region.

```{r, fig.width=6, fig.height=6}
iscam_normpower(LOS = .05, n = 20, prob1 = .5, alternative = "less", prob2 = .7)
```

## Probability Distributions
- **iscam_binom**

This function calculates either a tail probability or the inverse cumulative probability depending on whether k or prob is passed a question mark.

```{r, fig.width=6, fig.height=4}
iscam_binom("?", 0.05, 20, 0.3, lower.tail = TRUE)
iscam_binom(10, "?", 20, 0.3, TRUE)
```

- **iscam_binomprob**

This function calculates tail probabilities from the binomial distribution. 
```{r, fig.width=6, fig.height=4}
iscam_binomprob(k = 20, n = 30, prob = 0.5, lower.tail = TRUE)
```

- **iscam_hyperprob**

This function calculates tail probabilities from the hypergeometric distribution.

```{r, fig.width=6, fig.height=4}
iscam_hyperprob(k = 2, total = 52, succ = 5, n = 13, lower.tail = FALSE)
```

- **iscam_normprob**

This function calculates tail probability for the normal distribution. 

```{r, fig.width=6, fig.height=4}
iscam_normprob(xval = 2.2, direction = "above", label = "sample proportions")
```

- **iscam_chisqprob**

This function calculates the upper tail probability for the chi-square distribution.

```{r, fig.width=6, fig.height=4}
iscam_chisqprob(xval = 10, df = 4)
```

- **iscam_tprob**

This function calculates tail probability for the t distribution.

```{r, fig.width=6, fig.height=4}
iscam_tprob(xval = 4, df = 5, direction = "above")
```

## Normal Approximations
- **iscam_binomnorm**

This function illustrates the normal approximation to the binomial. 

```{r, fig.width=6, fig.height=4}
iscam_binomnorm(k = 20, n = 30, prob = 0.5, direction = "above")
```

- **iscam_hypernorm**

This function calculates tail probabilities from the hypergeometric distribution.

```{r, fig.width=6, fig.height=4}
iscam_hypernorm(k = 2, total = 52, succ = 5, n = 26, lower.tail = T)
```

## Histograms with Overlaid Density Curves
- **iscam_addexp**

This function creates a histogram of the inputted variable and overlays an exponential density function with lambda = 1/mean.

```{r, fig.width=6, fig.height=4}
expdata <- rexp(50)
iscam_addexp(expdata)
```

- **iscam_addlnorm**

This function creates a histogram of the inputted variable and overlays a log normal density function.

```{r, fig.width=6, fig.height=4}
lnormdata <- rlnorm(50)
iscam_addlnorm(lnormdata)
```

- **iscam_addnorm**

This function creates a histogram of the inputted variable and overlays a normal density function.

```{r, fig.width=6, fig.height=4}
normdata <- rnorm(50)
iscam_addnorm(normdata)
```

- **iscam_addt**

This function creates a histogram of the inputted variable and overlays a t density function with df degrees of freedom.

```{r, fig.width=6, fig.height=4}
tdata <- rt(50, df = 5)
iscam_addt(tdata, df = 5)
```

- **iscam_addtnorm**

This function creates a histogram of the inputted variable and overlays a t density function with df degrees of freedom and a normal density function.

```{r, fig.width=6, fig.height=4}
iscam_addtnorm(tdata, df = 3)
```